---
title: "Demographic Table"
author: Tingting Zhan
date: today
format: 
  html:
    page-layout: full
    html-math-method: katex
number-sections: true
toc: true
toc-location: left
toc-depth: 4
toc-title: ''
editor: source
knitr:
  opts_chunk: 
    collapse: true
    comment: "#" 
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---



# Introduction


```{r}
#| eval: false
#| echo: false
pak::pak('tingtingzhan/DemographicTable')
```

```{r}
#| eval: false
utils::install.packages('DemographicTable')
```

## Getting Started

Examples in this vignette requires

```{r}
library(DemographicTable)
```

Users may remove the last pipe `|> print()` from all examples if they are using R [`interactive`](https://search.r-project.org/R/refmans/base/html/interactive.html)ly.

# Demographic Table


## Summary of all subjects

```{r}
datasets::penguins |>
  subset(select = c('species', 'island', 'bill_len')) |>
  DemographicTable(data.name = 'datasets::penguins') |>
  print()
```

## Summary by one group

Color of each individual `group` is determined by `scales::pal_hue()`, which is the default color pallete used in package **`ggplot2`**.

```{r}
#| lst-label: lst-penguins_sex
#| lst-cap: '`penguins` by `sex`'
datasets::penguins |>
  subset(select = c('sex', 'species', 'bill_dep')) |>
  DemographicTable(by = ~ sex, data.name = 'datasets::penguins') |> 
  print()
```

User may choose to hide the $p$-values with option `compare = FALSE`.

```{r}
datasets::penguins |>
  subset(select = c('sex', 'species', 'bill_dep')) |>
  DemographicTable(by = ~ sex, data.name = 'datasets::penguins', compare = FALSE) |>
  print()
```

## Summary by multiple groups

```{r}
datasets::penguins |>
  subset(select = c('sex', 'island', 'species', 'bill_dep')) |>
  DemographicTable(by = ~ sex + island, data.name = 'datasets::penguins', compare = FALSE) |> 
  print()
```

# Combine Multiple `DemographicTable`s

```{r}
male = datasets::penguins |>
  subset(subset = (sex == 'male'), select = c('island', 'species', 'bill_dep')) |>
  DemographicTable(by = ~ island, data.name = 'Male Penguins', compare = FALSE)
```

```{r}
female = datasets::penguins |>
  subset(subset = (sex == 'female'), select = c('island', 'species', 'bill_dep')) |>
  DemographicTable(by = ~ island, data.name = 'Female Penguins', compare = FALSE)
```

```{r}
c(male, female) |> 
  print()
```

# Advance Usage

Remove the "overall" column.

```{r}
tb = datasets::penguins |>
  subset(select = c('sex', 'island', 'species', 'bill_dep')) |>
  DemographicTable(by = ~ sex + island, data.name = 'datasets::penguins', compare = FALSE)
tb[-1L] |> 
  print()
```


```{r}
c(male, female)[2:4] |> 
  print()
```

```{r}
c(male, female)[c(2L, 4L)] |> 
  print()
```

# Exception Handling

## Missing value in one or more groups

See @lst-penguins_sex.


## Use of `logical` values

Using [`logical`](https://search.r-project.org/R/refmans/base/html/logical.html) values is discouraged (@lst-logical), as this practice is proved confusing to scientists without a strong data background.

```{r}
#| warning: false
#| lst-label: lst-logical
#| lst-cap: 'Using `logical` values is discouraged'
datasets::mtcars |>
  within(expr = {
    vs_straight = as.logical(vs)
    am_manual = as.logical(am)
  }) |>
  subset(select = c('am_manual', 'drat', 'vs_straight')) |>
  DemographicTable(by = ~ am_manual, data.name = 'mtcars') |>
  print()
```


Instead of using [`logical`](https://search.r-project.org/R/refmans/base/html/logical.html) variables, we recommend using 2-`level` [`factor`](https://search.r-project.org/R/refmans/base/html/factor.html)s (@lst-factor).

```{r}
#| lst-label: lst-factor
#| lst-cap: 'We recommend using 2-`level` `factor`s'
datasets::mtcars |>
  within(expr = {
    vs = ifelse(vs, yes = 'Straight', no = 'V-shaped')
    am = ifelse(am, yes = 'manual', no = 'automatic')
  }) |> 
  subset(select = c('am', 'drat', 'vs')) |>
  DemographicTable(by = ~ am, data.name = 'mtcars') |>
  print()
```

